Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-01-29 colt-21/cvx-flats.tex
mostly intro

\usepackage{mathtools, verbatim}
%\usepackage[thmmarks, thref, amsthm]{ntheorem}
\usepackage{color}
\definecolor{darkblue}{rgb}{0.0,0.0,0.2}
\hypersetup{colorlinks,breaklinks,
	linkcolor=darkblue,urlcolor=darkblue,
	anchorcolor=darkblue,citecolor=darkblue}
\usepackage{wrapfig}
\usepackage{caption}
% NOTE: had to comment out "todo" in colt2021.cls
\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
%\usepackage{accents}
\usepackage{bbm}
\usepackage{xspace}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}

@@

\paragraph{The ``four quadrants'' of problem types}
Above, we discuss a significant divergence in previous frameworks: the difference in constructing a surrogate given a \emph{target loss} and a \emph{target statistic}.
In addition to the two possible targets, we may have one of two domains: a \emph{discrete} (i.e.\ finite) target prediction space, like a classification problem, or a \emph{continuous} one, like a regression or estimation problem.
We informally refer to the four resulting cases---target loss vs.\ target statistic, and discrete vs.\ continuous predictions---as the ``four quadrants'' of supervised learning problems, shown in Table~\ref{tab:quadrants}.
For further examples, see Appendix~\ref{app:omitted-examples}.

\paragraph{Literature on consistency and calibration}% \label{subsec:consistency}
%\bo{Citation help appreciated!} \jessie{\cite{fisher1922mathematical,zhang2004statistical,bartlett2006convexity,tewari2007consistency,steinwart2007compare,ramaswamy2016convex} -- these are the ones I know of that discuss consistency in slightly various forms.  I believe the definition we run with is most similar to \cite{ramaswamy2016convex}.}
%\jessiet{Make this subsection more concise?}
We focus on the construction of consistent surrogate losses $L: \reals^d \times \Y \to \reals$, roughly meaning that minimizing $L$-loss corresponds to solving the target problem of interest.
Typically, we are given a \emph{target loss} $\ell$, and we roughly define $L$ to be consistent if minimizing $L$ and applying a link, minimizes $\ell$ (Definition \ref{def:consistent-ell})~\citep{zhang2004statistical,bartlett2006convexity,tewari2007consistency,steinwart2007compare,ramaswamy2016convex}.
In Definition \ref{def:consistent-prop}, we introduce a notion of consistency in which one is given a \emph{target statistic}, such as the conditional variance or quantile, as in classical statistics, but no target loss~\citep{gyorfi2006distribution, fan1998efficient,ruppert1997local}.
Here we will define $L$ to be consistent if minimizing $L$ and applying a link function yields estimates converging to the correct value.

A priori, it is not clear that compatible definitions of consistency could be given for both target statistics and target losses.
We observe that, in fact, target losses are a special case of target statistics (\S~\ref{sec:consis-implies-indir}), which suggests property elicitation (see \S~\ref{subsec:properties}) is well-suited to study general lower bounds.
%Most prior work on convex consistency dimension focuses on problems in Quadrant 1~\citep{zhang2004statistical,bartlett2006convexity,tewari2007consistency,ramaswamy2015hierarchical,ramaswamy2016convex,ramaswamy2018consistent}.

@@

\tikzset{vertex/.style = {ellipse,draw,minimum width=10em, fill = white, font=\Large}}
\tikzset{edge/.style = {->,> = triangle 45, thick}}
\tikzset{node/.style = {anchor=above, sloped}}
\tikzstyle{every node}=[font=\Large]

@@

\node[vertex] (d-flats) at  (4,0) {$d$-flats in level sets};
\node[vertex, text width=45mm] (q1-bounds) at (12,2) {$\conscvx$ bounds for Q1,2. (Cor.~\ref{cor:fsd-bound})};
\node[vertex, text width=45mm] (q4-bounds) at (12,-2) {$\conscvx$ bounds for Q3,4. (Theorem~\ref{thm:bayes-risk-lower-bound})};

%edges
\draw[edge, <->] (consistent) to node[right,text width=5mm]{\cite{bartlett2006convexity}} (calibrated);
\draw[edge] (consistent) to node[above, sloped]{Theorem~\ref{thm:consistent-implies-indir-elic}} (indir-elic);
\draw[edge] (calibrated) to (indir-elic);
\draw[edge] (indir-elic) to node[above, sloped]{Lemma~\ref{lem:convex-flats-inf-dim}} (d-flats);

\draw[edge] (d-flats) to node[above, sloped]{Cor.~\ref{cor:Pcodim-flat-elic-relint-prop}} (q4-bounds);
\draw[edge] (d-flats) to node[above, sloped]{Cor.~\ref{cor:Pcodim-flat-single-val-prop}} (q1-bounds);
\end{tikzpicture}
}
\end{center}
\caption{Flow and implications of results \jessie{Change}\raf{Suggested tweaks: larger font, lower vertical distance, newlines as needed (see what I did in Bartlett et al.\ to make that happen)} \jessie{Attempted. Feel free to modify from here.}}
\label{fig:results-flow}
\end{figure}

\paragraph{Contributions}
\jessiet{Thoughts on this vs bullet points?}
First, we formalize a notion of consistency with respect to a target statistic (Definition~\ref{def:consistent-prop}) and show its relationship to consistency with respect to a target loss (Lemma~\ref{lem:consistent-loss-implies-prop}).
We then show indirect elicitation is a necessary condition for consistency (Theorem~\ref{thm:consistent-implies-indir-elic}).
With these tools in hand, we present a new framework for deriving lower bounds on the prediction dimension of consistent convex surrogates (Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}) via indirect elicitation.
These bounds are the first to our knowledge that can be applied in all four quadrants.
Finally, we use this framework to derive new bounds for well-studied problems such as abstain loss (\S~\ref{sec:finite-calib}) and variance, entropy, and norms (\S~\ref{sec:contin-consis}).
In Figure~\ref{fig:results-flow}, we show the implications of some of our main results.

@@

        \jessiet{Adding an attempt at first paragraph of sketch below, but also my brain isn't functioning.}
    If $\conscvx(\gamma) \geq d$, then there is a $L \in \Lcvx_d$ so that $L$ is consistent with respect to $\gamma$, and in turn, indirectly elicits $\gamma$.
    Lemma~\ref{lem:convex-flats-inf-dim} says that there is some $d$-flat $F = \zeros{V}$ contained in $\gamma_r$.
    In particular, we can verify $\Sc_{\gamma_r}(p) + \{p\}$ is one such $d$-flat.
    If $p \in \relint{\simplex}$, then the feasible subspace  (shifted by $p$) spans\jessiet{not the word I want...} $\relint{\simplex}$, and we can observe that the shifted feasible subspace is full dimensional relative to $\affhull(\simplex)$.
    We then use the rank-nullity theorem with a bound on $\rank(V) \leq d$ to obtain the bound $d \geq n-1-\dim(\Sc_{\gamma_r(p)})$.

>>>>

\usepackage{mathtools, verbatim}
%\usepackage[thmmarks, thref, amsthm]{ntheorem}
\usepackage{color}
\definecolor{darkblue}{rgb}{0.0,0.0,0.2}
\hypersetup{colorlinks,breaklinks,
	linkcolor=darkblue,urlcolor=darkblue,
	anchorcolor=darkblue,citecolor=darkblue}
\usepackage{wrapfig}
\usepackage[small]{caption}
% NOTE: had to comment out "todo" in colt2021.cls
\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
%\usepackage{accents}
\usepackage{bbm}
\usepackage{xspace}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}

@@

\paragraph{The ``four quadrants'' of problem types}
Above, we discuss a significant divergence in previous frameworks: constructing a surrogate given a \emph{target loss} versus a \emph{target statistic}.
In addition to the two possible targets, we may have one of two domains: a \emph{discrete} (i.e.\ finite) target prediction space, like a classification problem, or a \emph{continuous} one, like a regression or estimation problem.
We informally refer to the four resulting cases---target loss vs.\ target statistic, and discrete vs.\ continuous predictions---as the ``four quadrants'' of supervised learning problems, shown in Table~\ref{tab:quadrants}.
For further examples, see Appendix~\ref{app:omitted-examples}.

\paragraph{Literature on consistency and calibration}% \label{subsec:consistency}
%\bo{Citation help appreciated!} \jessie{\cite{fisher1922mathematical,zhang2004statistical,bartlett2006convexity,tewari2007consistency,steinwart2007compare,ramaswamy2016convex} -- these are the ones I know of that discuss consistency in slightly various forms.  I believe the definition we run with is most similar to \cite{ramaswamy2016convex}.}
%\jessiet{Make this subsection more concise?}
We focus on the construction of consistent surrogate losses $L: \reals^d \times \Y \to \reals$, roughly meaning that minimizing $L$-loss corresponds to solving the target problem of interest.
When given a \emph{target loss} $\ell$, we roughly define $L$ to be consistent if minimizing $L$, and applying a link function, minimizes $\ell$ (Definition \ref{def:consistent-ell})~\citep{zhang2004statistical,bartlett2006convexity,tewari2007consistency,steinwart2007compare,ramaswamy2016convex}.
When given a target statistic such as the conditional quantile or variance, but no target loss, we introduce a notion of consistency in line with classical statistics (Definition \ref{def:consistent-prop}) \citep{gyorfi2006distribution, fan1998efficient,ruppert1997local}.
Here we will define $L$ to be consistent if minimizing $L$ and applying a link function yields estimates converging to the correct value.

A priori, it is not clear that compatible definitions of consistency could be given for both target statistics and target losses.
In fact, we observe that consistency for target losses is a special case of consistency for target statistics (\S~\ref{sec:consis-implies-indir}).
This observation suggests property elicitation (see \S~\ref{subsec:properties}) as a useful tool to study general lower bounds.
%Most prior work on convex consistency dimension focuses on problems in Quadrant 1~\citep{zhang2004statistical,bartlett2006convexity,tewari2007consistency,ramaswamy2015hierarchical,ramaswamy2016convex,ramaswamy2018consistent}.

@@

\tikzset{vertex/.style = {draw,minimum width=13em, inner sep=8pt, fill = white, font=\Large}}
\tikzset{edge/.style = {->,> = triangle 45, thick}}
\tikzset{node/.style = {anchor=above, sloped}}
\tikzstyle{every node}=[font=\Large]

@@

\node[vertex] (d-flats) at  (4,0) {$d$-flats in level sets};
\node[vertex, text width=45mm] (q1-bounds) at (12,2) {$\conscvx$ bounds for Q1,2. (Cor.~\ref{cor:fsd-bound})};
\node[vertex, text width=45mm] (q4-bounds) at (12,-2) {$\conscvx$ bounds for Q3,4. (Thm.~\ref{thm:bayes-risk-lower-bound})};

%edges
\draw[edge, <->] (consistent) to node[right,text width=5mm]{\cite{bartlett2006convexity}} (calibrated);
\draw[edge] (consistent) to node[above, sloped]{Thm.~\ref{thm:consistent-implies-indir-elic}} (indir-elic);
\draw[edge] (calibrated) to (indir-elic);
\draw[edge] (indir-elic) to node[above, sloped]{Lem.~\ref{lem:convex-flats-inf-dim}} (d-flats);

\draw[edge] (d-flats) to node[above, sloped]{Cor.~\ref{cor:Pcodim-flat-elic-relint-prop}} (q4-bounds);
\draw[edge] (d-flats) to node[above, sloped]{Cor.~\ref{cor:Pcodim-flat-single-val-prop}} (q1-bounds);
\end{tikzpicture}
}
\end{center}
\caption{Flow and implications of our results. Compared to calibration, we suggest indirect elicitation as a simpler but almost-as-powerful necessary condition for consistency. In particular, we obtain a testable necessary condition, based on $d$-flats, for whether there exists a $d$-dimensional consistent convex surrogate. This condition recovers and strengthens existing calibration-based results.}
\label{fig:results-flow}
\end{figure}

\paragraph{Contributions}
First, we formalize a notion of consistency with respect to a target statistic (Definition~\ref{def:consistent-prop}) and show its relationship to consistency with respect to a target loss (Lemma~\ref{lem:consistent-loss-implies-prop}).
We then show indirect elicitation is a necessary condition for consistency (Theorem~\ref{thm:consistent-implies-indir-elic}).
With these tools in hand, we present a new framework for deriving lower bounds on the prediction dimension of consistent convex surrogates (Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}) via indirect elicitation.
These bounds are the first to our knowledge that can be applied in all four quadrants.
Moreover, our framework can also give tighter bounds than previously existed in the literature.
We illustrate this sharpness with new bounds for well-studied problems such as abstain loss (\S~\ref{sec:finite-calib}) and variance, CVaR, and other measures of risk and uncertainty (\S~\ref{sec:contin-consis}).
See Figure~\ref{fig:results-flow} for a roadmap of our main results.

@@

        \jessiet{Adding an attempt at first paragraph of sketch below, but also my brain isn't functioning.}
    If $\conscvx(\gamma) \leq d$, then there is a $L \in \Lcvx_d$ so that $L$ is consistent with respect to $\gamma$, and in turn, indirectly elicits $\gamma$.
    Lemma~\ref{lem:convex-flats-inf-dim} says that there is some $d$-flat $F = \zeros{V}$ contained in $\gamma_r$.
    In particular, if $p \in \relint{\simplex}$, we can verify $\spn(F) \subseteq \Sc_{\gamma_r}(p) + \{p\}$.
    The feasible subspace (shifted by $p$) spans\jessiet{not the word I want...} $\relint{\simplex}$, and we can observe that the shifted feasible subspace is full dimensional relative to $\affhull(\simplex)$.
    We then use the rank-nullity theorem with a bound on $\rank(V) \leq d$ to obtain the bound $d \geq n-1-\dim(\Sc_{\gamma_r(p)})$.
